sdmTMB: Fast, flexible, and user-friendly spatial and spatiotemporal generalized linear mixed-effects models

### 2021-12-22 doc/appendix-binomial-inla.Rmd
Copy edit appendices

---
title: "Appendix 4: species distribution model comparison between INLA and sdmTMB"
output:
  pdf_document:
    highlight: tango
---

@@

```{r inla}
tictoc::tic("INLA") # start timing

nyear <- length(unique(d$year))
iset <- inla.spde.make.index("i", n.spde = spde$n.spde, n.group = nyear)
A <- inla.spde.make.A(mesh = inla_mesh, loc = loc_xy, group = d$i_year)
sdat <- inla.stack(
  data = list(y = d$present),
  A = list(A, 1),
  effects = list(iset, year_factor = as.factor(d$year)),
  tag = "stdata"
)

>>>>

---
title: "Appendix 4: Species distribution model comparison between INLA and sdmTMB"
output:
  pdf_document:
    highlight: tango
---

@@

```{r inla}
tictoc::tic("INLA") # start timing
nyear <- length(unique(d$year))
iset <- inla.spde.make.index("i", n.spde = spde$n.spde, n.group = nyear)
A <- inla.spde.make.A(mesh = inla_mesh, loc = loc_xy, group = d$i_year)
sdat <- inla.stack(
  data = list(y = d$present),
  A = list(A, 1),
  effects = list(iset, year_factor = as.factor(d$year)),
  tag = "stdata"
)

### 2021-12-22 doc/appendix-pcod-VAST-tweedie.Rmd
Copy edit appendices

---
title: "Appendix 5: example of index standardization of fishery-independent survey data with VAST"
output:
  pdf_document:
    highlight: tango
---

>>>>

---
title: "Appendix 5: Comparison of index standardization of fishery-independent survey data with VAST and sdmTMB"
output:
  pdf_document:
    highlight: tango
---

### 2021-12-22 doc/appendix-snow-sdmTMB.Rmd
Copy edit appendices

---
title: "Appendix 3: example of a spatially varying effect of climate on citizen science count data with sdmTMB"
output:
  pdf_document:
    highlight: tango
bibliography: refs.bib
---

@@

```{r packages, message=FALSE, warning=FALSE, cache=FALSE}
library(ggplot2)
library(dplyr)
library(sdmTMB)
library(sf)
theme_set(ggsidekick::theme_sleek() + theme( # modifications for plotting sf data
  legend.position = "top", legend.justification = c(0, 0), # move legend to top left
  legend.title = element_text(size = 10, hjust = 0), legend.key.height = unit(0.3, "cm"), 
  panel.grid.major = element_line(colour = "grey90"), # for lat/lons
  axis.title = element_blank())) # lat/lon is clear from axis text
```

This illustrates the fitting of a model with a spatially varying coefficient for the effect a regional climate index on species count data using sdmTMB. 
We will use counts of Snow Owls (*Bubo scandiacus*) observed in North America during annual Christmas Bird Counts [CBCs, @cbc]. 
Each survey is conducted between December 14 to January 5 in 24.14 kilometre diameter circle. 
Circle locations are chosen by local volunteers and therefore reflect both locations of higher human densities and known bird watching locations.
Each circle (given a unique ID) is generally surveyed annually by at least 10 volunteers, or deemed inactive. 
While these locations are not random, the consistent sampling through time still make them useful for exploring spatiotemporal processes.

Snowy Owls (*Bubo scandiacus*) breed on the arctic tundra and are irruptive migrants appearing across mid latitudes of North America in winter.
The reasons for the interannual variation in the number of individuals migrating south are not well understood, but seem to be related to high abundances of food during the breeding season and therefore sharp increases in breeding ground population densities [@robillard2016]. 
The North Atlantic Oscillation Index (NAO) has been linked to productivity of both owls and their prey in Europe [@millon2014]. 
On average a positive NAO index results in colder and drier conditions in northern Canada and mild and wetter conditions in the eastern United States [@liu2021]. 
For this analysis, we have defined survey year as the year prior to/ending during the survey window. 
For these years, we have retrieved mean annual climate indices from the University of East Anglia's Climatic Research Unit ([ https://crudata.uea.ac.uk/cru/data/nao/](https://crudata.uea.ac.uk/cru/data/nao/)). 
Because both productivity and the choice of wintering location could be influenced by climate, we will test for a spatially variable effect of annual mean NAO index between 1979 and 2020 on winter owl abundance. 
Our model will span the southern boundary of their winter distribution in North America, the only portion of their range well surveyed by CBCs.

@@

```{r explore, fig.asp = 0.3, cache=FALSE}
ggplot(snow, aes(year, count, colour = nao)) + geom_jitter() + 
  scale_colour_viridis_c() + ggsidekick::theme_sleek()
```

@@

Our only fixed effect will be the NAO climate index (`nao`), but because of the wide spatial extent of these surveys, we also expect that the effect of NAO might not be consistent throughout North America. 
Therefore, effect of NAO will also be allowed to vary spatially using this formula: `spatial_varying = ~ 0 + nao`.
Because there does not appear to be much correlation between years, year will be included as a random factor: `(1|year_f)`. 
We will also include a fixed spatial random field (`spatial = "on"`) to account for consistent differences in habitat, proximity to the breeding grounds, and effects of flyways and barriers like coasts, and independent and identically distributed spatiotemporal random field (`spatiotemporal = "IID"`) to allow for annual deviations in these spatial effects due to factors like local weather and prey abundance. 
The random intercept for year allows the mean estimate for each of these spatiotemporal fields to vary. 
If years in sequence were strongly correlated with each other, we could use a random walk to link their mean estimates by setting `time_varying = ~ 1`.

```{r fit-poisson, echo=TRUE, eval=FALSE}
m0 <- sdmTMB(count ~ 1 + nao + (1|year_f),
             spatial_varying = ~ 0 + nao,
             time = "year",
             spatial = "on", 
             spatiotemporal = "IID",
             family = poisson(link = "log"),
             data = snow, mesh = mesh)
```

```{r fit-negb1, echo=TRUE, eval=FALSE}
m1 <- sdmTMB(count ~ 1 + nao + (1|year_f),
            spatial_varying = ~ 0 + nao,  
            time = "year",
            spatial = "on", 
            spatiotemporal = "IID",
            family = nbinom1(link = "log"),
            data = snow, mesh = mesh)
```

```{r fit-negb2, echo=TRUE, eval=FALSE}
m2 <- sdmTMB(count ~ 1 + nao + (1|year_f),
            spatial_varying = ~ 0 + nao,  
            time = "year",
            spatial = "on", 
            spatiotemporal = "IID",
            family = nbinom2(link = "log"),
            data = snow, mesh = mesh)
```

```{r save-models, warning=FALSE, message=FALSE, echo=FALSE, cache=FALSE, eval=FALSE}
saveRDS(m0, here::here("data/snow_w_main_effect_0_150km.rds"))
saveRDS(m1, here::here("data/snow_w_main_effect_1_150km.rds"))
saveRDS(m2, here::here("data/snow_w_main_effect_150km.rds"))
```

@@

```{r aic, echo=TRUE, cache=FALSE}
AIC(m0,m1,m2)
```

@@

```{r fit-negb2-reml, echo=TRUE, eval=FALSE}
m <- sdmTMB(count ~ 1 + nao + (1|year_f), 
            spatial_varying = ~ 0 + nao,
            time = "year", spatial = "on", spatiotemporal = "IID",
            family = nbinom2(link = "log"),
            reml = TRUE,
            data = snow, mesh = mesh)
```

```{r save-reml-model, warning=FALSE, message=FALSE, echo=FALSE, cache=FALSE, eval=FALSE}
saveRDS(m, here::here("data/snow_w_main_effect_150km_reml.rds"))
```

@@

<!-- \clearpage -->
We can inspect randomized quantile residuals, although this method is not guaranteed to be accurate.

@@

```{r shapes}
coast <- rnaturalearth::ne_coastline(scale = "medium", returnclass = "sf") %>% 
  sf::st_transform(crs = Albers) 
lakes <- sf::st_read(here::here("data/ne_10m_lakes"),  quiet = TRUE)
lakes <- lakes[lakes$scalerank==0,] %>% sf::st_transform(crs = Albers) # only large lakes
```

@@

```{r p-mean, message=FALSE, echo=TRUE, eval=TRUE, cache=FALSE}
b <- tidy(m)

p_mean <- p %>% group_by(CBCID) %>% summarise(
  X = mean(X), Y= mean(Y), 
  mean_est_count = exp(mean(est)), 
  nao_effect = exp(b[b$term == "nao", 2] + mean(zeta_s))
  )
```

```{r plot-map, echo=TRUE, eval=T, cache=T}
ggplot(data = p_proj) + geom_point(data = p_mean, 
     aes(X,Y, colour = nao_effect, size = mean_est_count), alpha = 0.5) +
  geom_sf(data = coast, colour = "gray50") +
  geom_sf(data = lakes, colour = "gray50", fill = NA) +
  coord_sf(xlim = c(min(p_proj$X)-50000, max(p_proj$X)-50000),# adjusts space on sides
           ylim = c(min(p_proj$Y), max(p_proj$Y))) +
  scale_colour_viridis_c(guide = guide_colourbar(
    direction = "horizontal", title.position = "top")) +
  guides(size = "none") + labs(colour = "NAO effect\non Snowy Owl count") +
  theme(legend.position = c(0.1,0.1), axis.title = element_blank())
```

@@

```{r plot-all-effects, echo=TRUE, eval=T, cache=T}
ggplot(data = filter(p_proj, year > 1978)) + 
  geom_point(data = filter(p, year > 1978), aes(X, Y, colour = exp(est)), 
             size = 0.4, alpha = 0.5) +
  coord_sf(xlim = c(min(p_proj$X)-50000, max(p_proj$X)-50000), 
           ylim = c(min(p_proj$Y), max(p_proj$Y)))+
  scale_colour_viridis_c(trans="sqrt", # makes variation in low numbers visible 
    # can also trim the extreme high values
    limits = c(0, quantile(exp(p$est), 0.995)), na.value = "yellow",
    guide = guide_colourbar(title.position = "top")) +
  labs(colour = "Predicted counts") + 
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  facet_wrap(~year_f)
```

@@

We can look at the spatial random effects (`omega_s`) that represent consistent deviations in space that are not accounted for by our fixed effects. 
In other words, these deviations represent consistent biotic and abiotic factors that are affecting winter owl counts.
These could be landcover, proximity to the breeding grounds, and effects of flyways and barriers like coasts.

```{r plot-spatial-effects, echo=TRUE, eval=T, cache=T}
ggplot(data = p_proj) + 
  geom_point(data = p, aes(X,Y, colour = omega_s), alpha = 0.5) +
  geom_sf(data = coast, colour = "gray50") +
  geom_sf(data = lakes, colour = "gray50", fill = NA) +
  coord_sf(xlim = c(min(p_proj$X)-50000, max(p_proj$X)-50000),
           ylim = c(min(p_proj$Y), max(p_proj$Y))) +
  scale_colour_viridis_c(guide = guide_colourbar(
    direction = "horizontal", title.position = "top")) +
  theme(legend.position = c(0.1,0.1), axis.title = element_blank())
```

@@

```{r plot-spatiotemporal-effects, echo=TRUE, eval=T, cache=T}
ggplot(data = filter(p_proj, year > 1978)) + 
  geom_point(data = filter(p, year > 1978), aes(X, Y, colour = epsilon_st), 
             size = 0.4, alpha = 0.5) +
  coord_sf(xlim = c(min(p_proj$X)-50000, max(p_proj$X)-50000), 
           ylim = c(min(p_proj$Y), max(p_proj$Y))) +
  scale_colour_viridis_c(guide = guide_colourbar(title.position = "top")) +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  facet_wrap(~year_f)
```

@@

In this case these only include the overall effect of mean annual NAO and the random intercepts for year.

```{r plot-fixed-effects, echo=TRUE, eval=T, cache=T}
ggplot(data = filter(p_proj, year > 1978)) + 
  geom_point(data = filter(p, year > 1978), aes(X, Y, colour = exp(est_non_rf)), 
             size = 0.4, alpha = 0.5) +
  coord_sf(xlim = c(min(p_proj$X)-50000, max(p_proj$X)-50000), 
           ylim = c(min(p_proj$Y), max(p_proj$Y)))+
  scale_colour_viridis_c(breaks = c(0.03, 0.05, 0.07), 
    guide = guide_colourbar(title.position = "top")) +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  facet_wrap(~year_f)
```

>>>>

---
title: "Appendix 3: Example of a spatially varying effect of climate on citizen science count data with sdmTMB"
output:
  pdf_document:
    highlight: tango
    toc: true
    number_sections: true
urlcolor: blue
highlight: "monochrome"
bibliography: refs.bib
---

@@

```{r packages, message=FALSE, warning=FALSE, cache=FALSE}
library(ggplot2)
library(dplyr)
library(sdmTMB)
library(sf)
```

```{r theme, echo=FALSE}
theme_set(theme_light() + theme( # modifications for plotting sf data
  legend.position = "top", legend.justification = c(0, 0), # move legend to top left
  legend.title = element_text(size = 10, hjust = 0), legend.key.height = unit(0.3, "cm"),
  panel.grid.major = element_line(colour = "grey90"), # for lat/lons
  axis.title = element_blank()
)) # lat/lon is clear from axis text
```

This appendix illustrates the fitting of a model with a spatially varying coefficient for the effect a regional climate index on species count data using sdmTMB. 
We will use counts of Snowy Owls (*Bubo scandiacus*) observed in North America during annual Christmas Bird Counts [CBCs, @cbc]. 
Each survey is conducted between December 14 to January 5 in 24.14 kilometre diameter circle. 
Circle locations are chosen by local volunteers and therefore reflect both locations of higher human densities and known bird watching locations.
Each circle (given a unique ID) is generally surveyed annually by at least 10 volunteers, or deemed inactive. 
While these locations are not random, the consistent sampling through time still make them useful for exploring spatiotemporal processes.

Snowy Owls (*Bubo scandiacus*) breed on the arctic tundra and are irruptive migrants appearing across mid latitudes of North America in winter.
The reasons for the interannual variation in the number of individuals migrating south are not well understood, but seem to be related to high abundances of food during the breeding season and therefore sharp increases in breeding ground population densities [@robillard2016]. 
The North Atlantic Oscillation Index (NAO) has been linked to productivity of both owls and their prey in Europe [@millon2014]. 
On average a positive NAO index results in colder and drier conditions in northern Canada and mild and wetter conditions in the eastern United States [@liu2021]. 
For this analysis, we have defined survey year as the year prior to/ending during the survey window. 
For these years, we have retrieved mean annual climate indices from the [University of East Anglia's Climatic Research Unit](https://crudata.uea.ac.uk/cru/data/nao/). 
Because both productivity and the choice of wintering location could be influenced by climate, we will test for a spatially variable effect of annual mean NAO index between 1979 and 2020 on winter owl abundance. 
Our model will span the southern boundary of their winter distribution in North America, the only portion of their range well surveyed by CBCs.

@@

```{r explore, fig.asp = 0.3, cache=FALSE}
ggplot(snow, aes(year, count, colour = nao)) +
  geom_jitter() +
  scale_colour_viridis_c() +
  theme_light()
```

@@

Our only fixed effect will be the NAO climate index (`nao`), but because of the wide spatial extent of these surveys, we also expect that the effect of NAO might not be consistent throughout North America. 
Therefore, effect of NAO will also be allowed to vary spatially using this formula: `spatial_varying = ~ 0 + nao`.
Because there does not appear to be much correlation between years, year will be included as a random factor: `(1 | year_f)`. 
We will also include a fixed spatial random field (`spatial = "on"`) to account for consistent differences in habitat, proximity to the breeding grounds, and effects of flyways and barriers like coasts, and independent and identically distributed spatiotemporal random field (`spatiotemporal = "iid"`) to allow for annual deviations in these spatial effects due to factors like local weather and prey abundance. 
The random intercept for year allows the mean estimate for each of these spatiotemporal fields to vary. 
If years in sequence were strongly correlated with each other, we could use a random walk to link their mean estimates by setting `time_varying = ~ 1`.
Since these models will take a while to fit, in practice, we might set `silent = FALSE` to monitor optimization progress.

```{r rds1, echo=FALSE, cache=FALSE}
rds_saved <- file.exists(here::here("data/snow_w_main_effect_0_150km.rds"))
```

```{r fit-poisson, echo=TRUE, eval=!rds_saved, warning=FALSE}
m0 <- sdmTMB(count ~ 1 + nao + (1 | year_f),
  spatial_varying = ~ 0 + nao,
  time = "year",
  spatial = "on",
  spatiotemporal = "IID",
  family = poisson(link = "log"),
  data = snow, mesh = mesh
)
```

```{r fit-negb1, echo=TRUE, eval=!rds_saved, warning=FALSE}
m1 <- sdmTMB(count ~ 1 + nao + (1 | year_f),
  spatial_varying = ~ 0 + nao,
  time = "year",
  spatial = "on",
  spatiotemporal = "IID",
  family = nbinom1(link = "log"),
  data = snow, mesh = mesh
)
```

```{r fit-negb2, echo=TRUE, eval=!rds_saved, warning=FALSE}
m2 <- sdmTMB(count ~ 1 + nao + (1 | year_f),
  spatial_varying = ~ 0 + nao,
  time = "year",
  spatial = "on",
  spatiotemporal = "IID",
  family = nbinom2(link = "log"),
  data = snow, mesh = mesh
)
```

```{r save-models, warning=FALSE, message=FALSE, echo=FALSE, cache=FALSE, eval=TRUE}
if (!rds_saved) {
  saveRDS(m0, here::here("data/snow_w_main_effect_0_150km.rds"))
  saveRDS(m1, here::here("data/snow_w_main_effect_1_150km.rds"))
  saveRDS(m2, here::here("data/snow_w_main_effect_150km.rds"))
}
```

@@

```{r aic, echo=TRUE, cache=FALSE}
AIC(m0, m1, m2)
```

@@

```{r rds2, echo=FALSE, cache=FALSE}
rds_saved2 <- file.exists(here::here("data/snow_w_main_effect_150km_reml.rds"))
```

```{r fit-negb2-reml, echo=TRUE, eval=!rds_saved2, warning=FALSE}
m <- sdmTMB(count ~ 1 + nao + (1 | year_f),
  spatial_varying = ~ 0 + nao,
  time = "year", 
  spatial = "on", 
  spatiotemporal = "iid",
  family = nbinom2(link = "log"),
  reml = TRUE,
  data = snow, mesh = mesh
)
```

```{r save-reml-model, warning=FALSE, message=FALSE, echo=FALSE, cache=FALSE, eval=FALSE}
if (!rds_saved2) {
  saveRDS(m, here::here("data/snow_w_main_effect_150km_reml.rds"))
}
```

@@

<!-- \clearpage -->
We can inspect randomized quantile residuals, although this method has known issues.

@@

```{r shapes}
coast <- rnaturalearth::ne_coastline(scale = "medium", returnclass = "sf") %>%
  sf::st_transform(crs = Albers)
lakes <- sf::st_read(here::here("data/ne_10m_lakes"), quiet = TRUE)
lakes <- lakes[lakes$scalerank == 0, ] %>% sf::st_transform(crs = Albers) # only large lakes
```

@@

```{r p-mean, message=FALSE, echo=TRUE, eval=TRUE, cache=FALSE}
b <- tidy(m)
p_mean <- p %>%
  group_by(CBCID) %>%
  summarise(
    X = mean(X), Y = mean(Y),
    mean_est_count = exp(mean(est)),
    nao_effect = exp(b[b$term == "nao", 2] + mean(zeta_s))
  )
```

```{r plot-map, echo=TRUE, eval=T, cache=T}
ggplot(data = p_proj) +
  geom_point(
    data = p_mean,
    aes(X, Y, colour = nao_effect, size = mean_est_count), alpha = 0.5
  ) +
  geom_sf(data = coast, colour = "gray50") +
  geom_sf(data = lakes, colour = "gray50", fill = NA) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000), # adjusts space on sides
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(guide = guide_colourbar(
    direction = "horizontal", title.position = "top"
  )) +
  guides(size = "none") +
  labs(colour = "NAO effect\non Snowy Owl count") +
  theme(legend.position = c(0.1, 0.1), axis.title = element_blank())
```

@@

```{r plot-all-effects, echo=TRUE, eval=T, cache=T}
ggplot(data = filter(p_proj, year > 1978)) +
  geom_point(
    data = filter(p, year > 1978), aes(X, Y, colour = exp(est)),
    size = 0.4, alpha = 0.5
  ) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000),
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(
    trans = "sqrt", # makes variation in low numbers visible
    # can also trim the extreme high values
    limits = c(0, quantile(exp(p$est), 0.995)), na.value = "yellow",
    guide = guide_colourbar(title.position = "top")
  ) +
  labs(colour = "Predicted counts") +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  facet_wrap(~year_f)
```

@@

We can look at the spatial random effects (`omega_s`) that represent consistent deviations in space that are not accounted for by our fixed effects. 
In other words, these deviations represent consistent biotic and abiotic factors that are affecting winter owl counts.
These could be landcover, proximity to the breeding grounds, and effects of flyways and barriers such as coasts.

```{r plot-spatial-effects, echo=TRUE, eval=T, cache=T}
ggplot(data = p_proj) +
  geom_point(data = p, aes(X, Y, colour = omega_s), alpha = 0.5) +
  geom_sf(data = coast, colour = "gray50") +
  geom_sf(data = lakes, colour = "gray50", fill = NA) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000),
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(guide = guide_colourbar(
    direction = "horizontal", title.position = "top"
  )) +
  theme(legend.position = c(0.1, 0.1), axis.title = element_blank())
```

@@

```{r plot-spatiotemporal-effects, echo=TRUE, eval=T, cache=T}
ggplot(data = filter(p_proj, year > 1978)) +
  geom_point(
    data = filter(p, year > 1978), aes(X, Y, colour = epsilon_st),
    size = 0.4, alpha = 0.5
  ) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000),
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(guide = guide_colourbar(title.position = "top")) +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  facet_wrap(~year_f)
```

@@

In this case, these only include the overall effect of mean annual NAO and the random intercepts for year.

```{r plot-fixed-effects, echo=TRUE, eval=T, cache=T}
ggplot(data = filter(p_proj, year > 1978)) +
  geom_point(
    data = filter(p, year > 1978), aes(X, Y, colour = exp(est_non_rf)),
    size = 0.4, alpha = 0.5
  ) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000),
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(
    breaks = c(0.03, 0.05, 0.07),
    guide = guide_colourbar(title.position = "top")
  ) +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  facet_wrap(~year_f)
```

### 2021-12-22 doc/paper.Rmd
Copy edit appendices

Appendix 1: sdmTMB statistical model description\
Appendix 2: Example of index standardization of fishery-independent survey data with sdmTMB\
Appendix 3: Example of a spatially varying effect of climate on citizen science count data with sdmTMB\
Appendix 4: Species distribution model comparison between INLA and sdmTMB\
Appendix 5: Example of index standardization of fishery-independent survey data with VAST\
Appendix 6: sdmTMB model validation and speed-testing methods
<!--
All submissions with more than one author must include an Authors’ contributions statement. All persons listed as authors on a paper are expected to meet ALL of the following criteria for authorship:

>>>>

Appendix 1: sdmTMB statistical model description\
Appendix 2: Example of index standardization of fishery-independent survey data with sdmTMB\
Appendix 3: Example of a spatially varying effect of climate on citizen science count data with sdmTMB\
Appendix 4: Species distribution model comparison between INLA and sdmTMB\
Appendix 5: Comparison of index standardization of fishery-independent survey data with VAST and sdmTMB\
Appendix 6: sdmTMB model validation and speed-testing methods
<!--
All submissions with more than one author must include an Authors’ contributions statement. All persons listed as authors on a paper are expected to meet ALL of the following criteria for authorship:
